A (sun)Shiny Recreation of the New York Times \n Weather Chart

### 2017-04-17 Paper.Rmd
interactivity and outcomes

Although we have a much smaller set of cities in our Shiny app, we wanted to include much of the same interactivity that the New York Times implemented, while also extending to add a few new ways that users could interact with the data. The main update we wanted to include was the ability to change which year the user wanted to have daily temperatures for. Although the original graphic allows for a lot of cool interpretability, the ability to compare across years is a huge advantage when investigating how weather trends have changed. Although we have daily temperature data ranging all the way back to 1945, the set of cities we are working with are inconsistent with their start dates. However, all cities in our data have daily temperatures beyond 1950, so we chose to make this element of the interactivity available for 1950 to 2016 only. Since the chart spans a full year, we do not include 2017 in the 'Single Year' tab of our Shiny app. Users can interact with data for the current year in the other tab, explained in section **something**.

As already discussed in the scraping section, we have a limited scope of cities. However, the interactivity with city is essentially the same as the New York Times chart, with the exception of the super cool moving globe thing. Users can choose a midwestern city from a drop-down menu that includes all available cities (that the app creators have scraped up to this point).

Users of the app can select which year they want to look at by moving a slider input across a variety of years. When users select a year, they are selecting which year they would like to see for the top layer. It is important to note that the record and historical averages layers will not change for any given city, no matter which year is selected. These are record values that are current through March 2017, either averaged with data from the current year (for average data), or updated when record data happens.

The most innovative update to the interactive version of this chart (or at least we think so) is that the user may choose which layers of the plot they would like to view together. A user can look at any of the three layers individually, or any combination of the three layers. This allows users to see the chart in all its glory, or just examine a single aspect of the data for that given city and year. The base layer and gridlines remain constant, and as users check boxes to select which layers they want to examine, the layers are added or removed from this base plot accordingly. No matter which order the boxes are checked in, layers will always be added in the order they are meant to be in the original chart (i.e., records on the bottom, then averages, and finally the daily temperatures on top). This interactivity was made possible by a reactive function that creates 7 plots of the possible layer combinations for the city-year pair, and returns the one that matches the boxes that the user has checked.

@@

@@

>>>>

The New York Times weather chart allows users to select a city from a drop down menu and view dail temperatures for the year 2015 along with recoreds and averages. We wanted to keep this useful peice of interactivity and so our users can too chose a city to view from a drop down menu. As already discussed in the scraping section, we have a limited scope of cities. When a user choses one of the midwestern cities from the menu, the data is filtered to include temperatures for that city only.

Next, we wanted to implement an interactive option that wasn't available on the NYT weather chart. We thought that by allowing users to view daily temperatures for a large range of years, not just for 2015 would be a large improvement. Some of the cities that we scraped data for have daily temperature data ranging all the way back to 1945, but unfortunately the full set of cities we are working with do not agree on this start date. All of the cities do have daily temperatures beyond 1950, so we chose to make this element of the interactivity available to users only for the years 1950 to 2016 in the form of a slider bar. Since the chart spans a full year, we do not include 2017 in the 'Single Year' tab of our Shiny app. Users can view data for the current year in the other tab, explained in Section 2.3.2.

When users select a single year, they are dictating the set of daily temperatures that will be filtered from the data and plotted in the top layer of the graphic. It is important to note that the record and historical averages layers will not change for any given city, no matter which year is selected. These are record values that are current through March 2017, either averaged with data from the current year (for average data), or updated when a record temperature occurs.

Additionally, we decided to allow the user to select which layers of the visual they would like to view together. The base layer and gridlines will remain constant, and as users check boxes to select which layers they want to examine, the layers are added or removed from this base plot accordingly. No matter which order the boxes are checked in, layers will always be added in the order they are meant to be in the original chart (i.e., records on the bottom, then averages, and finally the daily temperatures on top). This interactivity was made possible by a reactive function that creates seven plots of each possible layer combination for the city-year pair, and returns the one that matches the boxes that the user has checked.

@@

While the original chart recreation in the 'Single Year' tab is very informative and flexible, it only allows users to inspect a single year at a time. We thought it would be nice if the user could compare temperature data for a single city over many years in one place so we created a spaghetti plot feature that resides in the 'Spaghetti Plot' tab of the Shiny app.

Again, users are able to choose a city to investigate from a drop down menu. For the chosen city, the spaghetti plot is constructed with temperatures on the y-axis and month-day on the x-axis. Data is added to the chart by plotting a line through the average daily temperatures over the course of one year for each year. Lines for the years 1950 through 2015 are grey while the lines for 2016 and 2017 are colored so that the user is able to quickly see recent weather patterns and how they relate to previous years for a single city.

@@

In the 'Single Year' tab of the Shiny app, users can interact with our recreation of the original chart. Although we have a much smaller set of cities in our Shiny app, we wanted to include much of the same interactivity that the New York Times implemented, while also extending to add a few new ways that users could interact with the data. The largest and most straight forward update that we wanted to include was the ability for users to choose which year to display daily temperatures for. The original graphic has many other impressive details, but the ability to view temperatures for a range of years will be a great advantage when investigating how weather trends change over time. Users of the app can select which year they want to look at by moving a slider input across a variety of years and the app reacts accordinly to change the top layer to reflect current temperatures for the chosen year.

What we might deem the most innovative update to the interactive version of this chart is the flexibility and variety of viewing options that are available to the user. We allow the user to look at any of the three layers individually, or any combination of the three layers. This allows them to see the full graphic that resembles the original New York Times chart, or just examine a single aspect of the data for a specified city and year.

In the 'Spaghetti Plot' tab the user can view average trends for a single city over multiple years in one graphic. This is advantagous to a user who is interested in assessing and comparing average weather trends between many years. Additionally this plot includes data for the year 2017, which the single year plot does not. In order to show the user what average temperatures have been doing lately, we highlight the latest complete year (2016) and the current incomplete year.
